---
title: "Evaluation of Motorcar Transmissions"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Executive Summary

This work is intended to answer the following questions:
* Is an automatic or manual transmission better for MPG?
* Quantify the MPG difference between automatic and manual transmissions.

### Results


## Data Preparation & Cleaning

Base dataset mtcars does not have factors by default, so I'll create them here.

```{r dataprep}
data(mtcars)
mtcars$cyl  <- factor(mtcars$cyl)
mtcars$vs   <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am   <- factor(mtcars$am, labels=c('Auto', 'Manual'))
```

## Exploratory Data Analysis

First, I explore the variables and their relationship with a pairs plot, stored 
in \textbf{Appendix-Figure-1}. The variables \textcolor{red}{cyl}, \textcolor{red}{disp},
\textcolor{red}{hp}, \textcolor{red}{drat}, \textcolor{red}{wt}, \textcolor{red}{vs},
\textcolor{red}{am}, and \textcolor{red}{carb} seem to have reasonable correlation with \textcolor{red}{mpg}, but
the linear regressions will be the real check of that. However, a box-plot may 
also be of value to compare \textcolor{red}{mpg} between 'Auto' vs 'Manual' in 
\textcolor{red}{am}. See \textbf{Appendix-Figure-2} for this plot.

## Model Selection

```{r appendix1, echo = FALSE, fig.width = 10, fig.height = 10, fig.cap = 'Appendix-Figure-1'}

# Taken from http://stackoverflow.com/questions/15271103/how-to-modify-this-correlation-matrix-plot
panel.cor <- function(x, y, digits = 2, cex.cor)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  test <- cor.test(x,y)
  Signif <- ifelse(round(test$p.value,3)<0.001,"p<0.001",paste("p=",round(test$p.value,3)))  
  text(0.5, 0.25, paste("r=",txt))
  text(.5, .75, Signif)
}

panel.smooth<-function (x, y, col = "blue", bg = NA, pch = 18, 
                        cex = 0.8, col.smooth = "red", span = 2/3, iter = 3, ...) 
{
  points(x, y, pch = pch, col = col, bg = bg, cex = cex)
  ok <- is.finite(x) & is.finite(y)
  if (any(ok)) 
    lines(stats::lowess(x[ok], y[ok], f = span, iter = iter), 
          col = col.smooth, ...)
}

panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col="cyan", ...)
}

pairs(mtcars, 
      diag.panel=panel.hist,
      gap = .5,
      main = "mtcars Variables Ordered and Colored by Correlation",
      lower.panel = panel.smooth,
      upper.panel = panel.cor)
```

```{r appendix2, echo = FALSE, fig.width = 10, fig.height = 6, fig.cap = 'Appendix-Figure-2'}
ggplot(data = mtcars, aes(x = am, y = mpg, group = am, fill = am)) +
         geom_boxplot(size = 1) +
         theme_bw() +
         ggtitle("mtcars MPG (Miles per Gallon) by Transmission Type (Auto vs Manual)") +
         labs(x = "Transmission (Auto vs Manual)", y = "MPG (Miles per Gallon)") +
         theme(legend.position = "none") +
        scale_fill_manual(values = c("#ff0000", "#0066ff"))
```


## Additional Resources

* [Github Repository](https://github.com/bkkesseler/DSS7-Regression-Models-Week4)

## Session Information

- 3.20 GHz Intel i5 650
- 8GB RAM
- RStudio Version 0.99.902

```{r sessioninfo}
sessionInfo()
```